Unimodular covers of 3-dimensional parallelepipeds and Cayley sums

### 2019-07-02 unimodular-covers.tex
expanding proofs in appendix

\begin{proof}
Suppose by contradiction that there is no lattice point as described in the lemma. Then there cannot be a vertex of $P$ in the corresponding strip, since $P$ is a lattice polygon. The same holds for $Q$ and its corresponding strip. Thus $P$ has two edges which each have one vertex on each side of the strip $(p_1p_2 + \langle \overrightarrow{q_1q_2}\rangle)$; we will call these edges $e_1$ and $e_2$. The same holds for $Q$ and the strip $(q_1q_2 + \langle \overrightarrow{p_1p_2}\rangle)$, and we call the edges be $f_1, f_2$.

@@

claim: if we tilt on of the $f_i$s so that they are parallel, there is no room to fit $e_i$, since it intersects both.

@@

\begin{proof}
$l_2, r_2, b_1, t_1 \ge 1$ follows from the fact that each edge crosses the corresponding strip. If one of the inequalities is not strict then one of the strips contains half of the square $p_1p_2+q_1q_2$, hence it contains a lattice point in its interior.  
\end{proof}

@@

\begin{proof}
If, say, $b_2 \le 0 \le t_2$. Then the intersections of $t$ and $b$ with the vertical line $\{x=-\beta\}$ through $q_1$ are less than distance one apart (because $t$ must go below every lattice point in the interior of $p_1p_2 + q_1q_2 - p_1$ and $b$ above every latttice point in the interior of $p_1p_2 + q_1q_2 - p_2$). This implies the vertical distance between any two points in $Q \cap \{x\le  -\beta\}$ is less than one, so it is impossible to fit a translated copy of $l$ in it.
\end{proof}

@@

\begin{proof}
The assumption implies that the intersections of $t$ and $b$ with the vertical line $\{x=1-\beta\}$ through $q_2$ are less than distance one apart (consider a vertical lattice line intersecting $q_1q_2$ and two consecutive lattice points in it, one above and one below the horizontal axis). This in turn implies part (a). For (b), if $r_1$ was $\le 0$ it would be impossible to fit a translated copy of $r$ in the right side of $Q$, because of (a).
\end{proof}

>>>>

\begin{proof}
%Suppose by contradiction that there is no lattice point as described in the lemma. Then there cannot be a vertex of $P$ in the corresponding strip, since $P$ is a lattice polygon. The same holds for $Q$ and its corresponding strip. Thus $P$ has two edges which each have one vertex on each side of the strip $(p_1p_2 + \langle \overrightarrow{q_1q_2}\rangle)$; we will call these edges $e_1$ and $e_2$. The same holds for $Q$ and the strip $(q_1q_2 + \langle \overrightarrow{p_1p_2}\rangle)$, and we call the edges be $f_1, f_2$.

Suppose by contradiction that there is no lattice point as described in the lemma. Then there cannot be a vertex of $P$ in the corresponding strip, since $P$ is a lattice polygon. The same holds for $Q$ and its corresponding strip. Thus $P$ has two primitive edges which each have one vertex on each side of the strip $p_1p_2 + \langle \overrightarrow{q_1q_2}\rangle$; we will call these edges $\ell$ and $r$ (with $\ell$ to the left and $r$ to the right). The same holds for $Q$ and the strip $q_1q_2 + \langle \overrightarrow{p_1p_2}\rangle$, and we call the edges  $b, t$ (with $b$ below and $t$ above). To make things more explicit, we assume the following coordinates for the edge-vectors:
\[
\vec l = (l_1,l_2), \quad
\vec r= (r_1,r_2), \quad
\vec b= (b_1,b_2), \quad
\vec t= (t_1,t_2).
\]
Without loss of generality we assume $l_2,r_2,b_1,t_1 >0$. (They are non-zero and we can arbitrarily choose a sign for each vector).

@@

claim: if we tilt one of the $f_i$s so that they are parallel, there is no room to fit $e_i$, since it intersects both.

@@

\begin{proof}
$l_2, r_2, b_1, t_1 \ge 1$ follows from the fact that each edge crosses the corresponding strip. 
If one of the inequalities is not strict, then one edge must have vertices on the corresponding strip. Then one of $(p_1p_2 + \langle \overrightarrow{q_1q_2}\rangle) \cap P$ or $(q_1q_2 + \langle \overrightarrow{p_1p_2}\rangle) \cap Q$
contains a triangle with one edge a translation of $p_1p_2$ and another edge a translation of $q_1q_2$. Such a triangle must contain a lattice point other than its vertices, and since $p_1p_2$ and $q_1q_2$ are primitive, the lattice point must be in the interior of the strip.
\end{proof}

@@

\begin{proof}
If, say, $b_2 \le 0 \le t_2$. Then the intersections of $t$ and $b$ with the vertical line $\{x=-\beta\}$ through $q_1$ are less than distance one apart (because $t$ must go below every lattice point in the interior of $p_1p_2 + q_1q_2 - p_1$ and $b$ above every latttice point in the interior of $p_1p_2 + q_1q_2 - p_2$). This implies the width w.r.t to $f_q$, the functional constant on $q_1q_2$, of any segment with vertices in $Q \cap \{x\le  -\beta\}$ is less than one, so it is impossible to fit a translated copy of $l$ in it.
\end{proof}

@@

\begin{proof}
The assumption implies that the intersections of $t$ and $b$ with the vertical line $\{x=1-\beta\}$ through $q_2$ are less than distance one apart (consider a vertical lattice line intersecting $q_1q_2$ and two consecutive lattice points in it, one above and one below the horizontal axis). This in turn implies part (a).

For (b), if $r_1$ was $\le 0$ it would be impossible to fit a translated copy of $r$ in the right side of $Q$, because of (a).
\end{proof}
